Gentoo Package Dependencies over Time

### 2014-08-08 Paper/paper.tex
Rounded results
Checkin

\usepackage{amsmath}
\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{microtype} % Finetunes spacing, load after fonts
\usepackage{siunitx}
\usepackage{alltt}
\usepackage{wrapfig}
\usepackage[colorlinks=false, pdfborder={0 0 0}]{hyperref} % Should be loaded last
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{subfigure}
\newcommand{\nd}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\tn}{\newline}

@@

\begin{verbatim}
cvs -d :pserver:anonymous@anoncvs.gentoo.org\
  /var/cvsroot co -D 12/01/2003 gentoo-x86
\end{verbatim}

@@

\begin{figure}[t]
\small\centering
\begin{alltt}
>=media-libs/taglib-1.6.1[asf,mp4]
>=media-libs/taglib-extras-1.0.1
player? (
    app-crypt/qca:2
    >=app-misc/strigi-0.5.7[dbus,qt4]
    || ( >=dev-db/mysql-5.0.76
         =virtual/mysql-5.1 )
    >=kde-base/kdelibs-4.3[opengl?,semantic-desktop?]
    sys-libs/zlib
    x11-libs/qt-script
    >=x11-libs/qtscriptgenerator-0.1.0
\end{alltt}
\vspace{-1em}
\caption{Fragment of the runtime dependencies of the Amarok music player.}\label{fig:amarokruntime}
\vspace{-2em}
\end{figure}

@@

\begin{wraptable}{r}{0.5\textwidth}
\vspace{-1em}
\small\centering
\caption{Key figures of the dataset}\label{tbl:dataset}
\begin{tabular}{ll}
\toprule
\multicolumn{2}{c}{\emph{Time}} \\[0.5mm]
Period & 2000 to Q1 2012 \\
Resolution & Monthly \\
Snapshots & 154 Graphs \\
\midrule
\multicolumn{2}{c}{\emph{All graphs}} \\[0.5mm]
Nodes & 1.3 Million \\
Vertices & 6.9 Million \\
\midrule
\multicolumn{2}{c}{\emph{Final graph}} \\[0.5mm]
Nodes & 15 Thousand \\
Vertices & 80 Thousand \\
\midrule
\multicolumn{2}{c}{\emph{Raw ebuilds}} \\[0.5mm]
Format & SquashFS \\
Size & 239 MB (compressed) \\
 & 4.4 GB (uncompressed) \\
Files & 2,990,722 \\
\midrule
\multicolumn{2}{c}{\emph{Compressed graphs}} \\[0.5mm]
Format & Tar+XZ compressed GML \\
Size & 2 MB \\
\bottomrule
\end{tabular}
\end{wraptable}

@@

The dataset collected is available at
\\\url{http://datahub.io/dataset/gentoo-dependency-graph}\\
as compressed GML graphs for every snapshot. Raw data is available at
\\\texttt{http://www.utwente.nl/mb/iebis/staff/amrit/}\\\texttt{timeline.squashfs}\\
as a SquashFS compressed filesystem. The SquashFS format allows one to mount the compressed data as a read-only disk drive and operate directly on the dataset. It contains the pre-processed ebuild files and the file \verb|graphs.bip| that contains all of the graphs in a custom format specialized for quick processing of large graphs over time. A C++ toolkit to work with the \verb|graphs.bip| file is available at \url{https://github.com/Recmo/depgraph}.

@@

Since it was impossible to get a visual overview of the entire dependency graph, its structure was plotted using double logarithmic histograms. It was found that there there are many packages with only zero, one or a few dependencies and a few packages with a lot of dependencies. Statistically this means the distribution of the number of dependencies has a fat tail. Likewise, the number of dependers for each package was plotted, this can be interpreted as the distribution of the number of adopters of a given technology. Again, there was a fat tail, even fatter than the one from the number of dependencies. The histogram was approximately linear in the double logarithmic scale, suggesting a power-law like distribution of the number of adopters for a given technology. \citet{zheng08} suggest that the structure of package dependency networks is not comparable to known models of social networks and have developed their own model to explain the graph structure.

@@

\section{Results}

@@

\begin{figure}
\small\centering
\includegraphics[width=\linewidth]{BassFit-git2.pdf}
\vspace{-2em}
\caption{The imitation driven adoption of \texttt{git}.}\label{fig:git}
\vspace{-1.5em}
\end{figure}

@@

@@

\begin{figure}
\small\centering
\includegraphics[width=\linewidth]{BassFit-libmad2.pdf}
\vspace{-2em}
\caption{The innovation driven adoption of \texttt{libmad}.}\label{fig:libmad}
\vspace{-1em}
\end{figure}

@@

\begin{figure}
\small\centering
\includegraphics[width=\linewidth]{BassFit-xulrunner-22.pdf}
\vspace{-2em}
\caption{The rise and decline of \texttt{xulrunner}.}\label{fig:xulrunner}
\vspace{-1em}
\end{figure}

@@

\subsection{Moooore}

\begin{figure}
\small\centering
\caption[Dependee growth after package introduction]{The Bass curve fitted on a number of packages}
\label{fig:BassFits}
\subfigure[\texttt{libnotify}: $p=0.05$, $q=0.72$]{\includegraphics[width=0.45\textwidth]{BassFit-libnotify.pdf}}
\subfigure[\texttt{libtheora}: $p=0.11$, $q=0.63$]{\includegraphics[width=0.45\textwidth]{BassFit-libtheora.pdf}}
\subfigure[\texttt{taglib}: $p=0.22$, $q=0.04$]{\includegraphics[width=0.45\textwidth]{BassFit-taglib.pdf}}
\subfigure[\texttt{udev}: $p=0.01$, $q=0.50$]{\includegraphics[width=0.45\textwidth]{BassFit-udev.pdf}}
\subfigure[\texttt{cairo}: $p=0.005$, $q=0.35$]{\includegraphics[width=0.45\textwidth]{BassFit-cairo.pdf}}
\end{figure}

@@

\begin{table}
\small\centering
\caption{Results of fitting the Bass model.}\label{tbl:results}
\begin{tabular}{lr@{ $\pm$}rr@{ $\pm$}rr@{ $\pm$}rr}
\toprule
Package & \multicolumn{2}{c}{$p$} & \multicolumn{2}{c}{$q$} & \multicolumn{2}{c}{$M$} & $\overline{R}^2$\\
\midrule
% 34, 65, 218, 77, 157, 240, 242, 17, 206, 143
\texttt{dev-dotnet/gconf-sharp}   & 0.167587 & 0.211648 & 0.754343 & 0.433141 & 21.283 & 0.484352 & 99.6148\%\\
\texttt{dev-haskell/cabal}        & 0.0554162 & 0.0732542 & 0.153744 & 0.379092 & 245.886 & 437.01 & 98.5521\%\\
\texttt{dev-libs/libunique}       & 0.0473548 & 0.0983485 & 0.136046 & 0.401718 & 286.948 & 731.387 & 98.5822\%\\
\texttt{dev-python/lxml}          & 0.00673774 & 0.00829268 & 0.48833 & 0.216921 & 52.883 & 52.4726 & 98.8237\%\\
\texttt{dev-python/simplejson}    & 0.0108121 & 0.0283649 & 0.463439 & 0.458178 & 97.1942 & 147.927 & 98.9801\%\\
\texttt{dev-vcs/git}              & 0.000415898 & 0.00512837 & 0.840506 & 0.143992 & 734.243 & 368.631 & 99.2386\%\\
\texttt{dev-vcs/mercurial}        & 0.00103413 & 0.00494549 & 0.516704 & 0.258352 & 173.8 & 480.637 & 98.2389\%\\
\texttt{gnome-base/gnome-keyring} & 0.0259038 & 0.0356738 & 0.833393 & 0.132449 & 41.5442 & 1.06264 & 99.6646\%\\
\texttt{media-libs/libcanberra}   & 0.0551051 & 0.116892 & 1.43593 & 0.438961 & 37.456 & 2.37536 & 99.6823\%\\
\texttt{x11-misc/xdg-utils}       & 0.0496807 & 0.0427185 & 0.257095 & 0.241258 & 114.925 & 126.964 & 99.5608\%\\
\bottomrule
\end{tabular}
\end{table}

@@

\section{Implications to Practice and Research}

>>>>

\usepackage{amsmath}
\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{microtype} % Finetunes spacing, load after fonts
\usepackage{siunitx}
\usepackage{alltt}
\usepackage{fancyvrb}
\usepackage{wrapfig}
\usepackage[colorlinks=false, pdfborder={0 0 0}]{hyperref} % Should be loaded last
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{subfigure}
\newcommand{\nd}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\tn}{\newline}

@@

\begin{center}
\begin{BVerbatim}
cvs -d :pserver:anonymous@anoncvs.gentoo.org\
  /var/cvsroot co -D 12/01/2003 gentoo-x86
\end{BVerbatim}
\end{center}

@@

\begin{figure}[t]
\small\centering
\begin{BVerbatim}
>=media-libs/taglib-1.6.1[asf,mp4]
>=media-libs/taglib-extras-1.0.1
player? (
    app-crypt/qca:2
    >=app-misc/strigi-0.5.7[dbus,qt4]
    || ( >=dev-db/mysql-5.0.76
         =virtual/mysql-5.1 )
    >=kde-base/kdelibs-4.3[opengl?,semantic-desktop?]
    sys-libs/zlib
    x11-libs/qt-script
    >=x11-libs/qtscriptgenerator-0.1.0
\end{BVerbatim}
\caption{Fragment of the runtime dependencies of the Amarok music player.}\label{fig:amarokruntime}
\end{figure}

@@

\begin{wraptable}{r}{0.5\textwidth}
\vspace{-1em}
\small\centering
\caption{Key figures of the dataset}\label{tbl:dataset}
\begin{tabular}{ll}
\toprule
\multicolumn{2}{c}{\emph{Time}} \\[0.5mm]
Period & 2000 to Q1 2012 \\
Resolution & Monthly \\
Snapshots & 154 Graphs \\
\midrule
\multicolumn{2}{c}{\emph{All graphs}} \\[0.5mm]
Nodes & 1.3 Million \\
Vertices & 6.9 Million \\
\midrule
\multicolumn{2}{c}{\emph{Final graph}} \\[0.5mm]
Nodes & 15 Thousand \\
Vertices & 80 Thousand \\
\midrule
\multicolumn{2}{c}{\emph{Raw ebuilds}} \\[0.5mm]
Format & SquashFS \\
Size & 239 MB (compressed) \\
 & 4.4 GB (uncompressed) \\
Files & 2,990,722 \\
\midrule
\multicolumn{2}{c}{\emph{Compressed graphs}} \\[0.5mm]
Format & Tar+XZ compressed GML \\
Size & 2 MB \\
\bottomrule
\end{tabular}
\vspace{-1em}
\end{wraptable}

@@

The dataset collected is available at
\begin{center}
\href{http://datahub.io/dataset/gentoo-dependency-graph}{\texttt{http://datahub.io/dataset/gentoo-dependency-graph}}
\end{center}
as compressed GML graphs for every snapshot. Raw data is available at
\begin{center}
\url{http://www.utwente.nl/mb/iebis/staff/amrit/timeline.squashfs}\\
\end{center}
as a SquashFS compressed filesystem. The SquashFS format allows one to mount the compressed data as a read-only disk drive and operate directly on the dataset. It contains the pre-processed ebuild files and the file \verb|graphs.bip| that contains all of the graphs in a custom format specialized for quick processing of large graphs over time. A C++ toolkit to work with the \verb|graphs.bip| file is available at \url{https://github.com/Recmo/depgraph}.

@@

@@

\section{Results and Discussion}

% \begin{wrapfigure}{r}{.5\linewidth}
% \small\centering
% \label{fig:BassFits}
% \subfigure[\texttt{libnotify}: $p=0.05$, $q=0.72$]{\includegraphics[width=0.45\textwidth]{BassFit-libnotify.pdf}}
% \subfigure[\texttt{libtheora}: $p=0.11$, $q=0.63$]{\includegraphics[width=0.45\textwidth]{BassFit-libtheora.pdf}}
% \subfigure[\texttt{taglib}: $p=0.22$, $q=0.04$]{\includegraphics[width=0.45\textwidth]{BassFit-taglib.pdf}}
% \subfigure[\texttt{udev}: $p=0.01$, $q=0.50$]{\includegraphics[width=0.45\textwidth]{BassFit-udev.pdf}}
% \subfigure[\texttt{cairo}: $p=0.005$, $q=0.35$]{\includegraphics[width=0.45\textwidth]{BassFit-cairo.pdf}}
% \caption{The Bass curve fitted on a number of packages}
% \end{wrapfigure}

@@

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\small\centering
\includegraphics[width=\linewidth]{BassFit-git2.pdf}
\caption{The imitation driven adoption of \texttt{git}.}\label{fig:git}
\vspace{-1em}
\end{wrapfigure}

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-2em}
\small\centering
\includegraphics[width=\linewidth]{BassFit-libmad2.pdf}
\caption{The innovation driven adoption of \texttt{libmad}.}\label{fig:libmad}
\vspace{-1em}
\end{wrapfigure}

@@

\begin{figure}
\small\centering
\includegraphics[width=0.6\linewidth]{BassFit-xulrunner-22.pdf}
\caption{The rise and decline of \texttt{xulrunner}. The model is fitted to the data up to 2011 (black dots).}\label{fig:xulrunner}
\end{figure}

@@

@@

\begin{table}
\small\centering
\caption{Results of fitting the Bass model.}\label{tbl:results}
\begin{tabular}{lr@{ $\pm$}rr@{ $\pm$}rr@{ $\pm$}rr}
\toprule
Package & \multicolumn{2}{c}{$p$} & \multicolumn{2}{c}{$q$} & \multicolumn{2}{c}{$M$} & $\overline{R}^2$\\
\midrule
% 34, 65, 218, 77, 157, 240, 242, 17, 206, 143
\texttt{dev-dotnet/gconf-sharp}   & 0.168 & 0.212 & 0.754 & 0.434 &  21 &   1 & 99.61\%\\
\texttt{dev-haskell/cabal}        & 0.055 & 0.074 & 0.154 & 0.380 & 246 & 438 & 98.55\%\\
\texttt{dev-libs/libunique}       & 0.047 & 0.099 & 0.136 & 0.402 & 287 & 732 & 98.58\%\\
\texttt{dev-python/lxml}          & 0.007 & 0.009 & 0.488 & 0.217 &  53 &  53 & 98.82\%\\
\texttt{dev-python/simplejson}    & 0.011 & 0.029 & 0.463 & 0.459 &  97 & 148 & 98.98\%\\
\texttt{dev-vcs/git}              & 0.000 & 0.006 & 0.841 & 0.144 & 734 & 369 & 99.24\%\\
\texttt{dev-vcs/mercurial}        & 0.001 & 0.005 & 0.517 & 0.259 & 174 & 481 & 98.24\%\\
\texttt{gnome-base/gnome-keyring} & 0.026 & 0.036 & 0.833 & 0.133 &  42 &   2 & 99.66\%\\
\texttt{media-libs/libcanberra}   & 0.055 & 0.117 & 1.436 & 0.439 &  37 &   3 & 99.68\%\\
\texttt{x11-misc/xdg-utils}       & 0.050 & 0.043 & 0.257 & 0.242 & 115 & 127 & 99.56\%\\
\bottomrule
\end{tabular}
\end{table}

% TODO: Write something about discretization noise

% TODO:

@@

\subsection{Limitations}

\subsection{Implications to Research and Practice}
